### 2015-05-26 conclusion.tex
Results

\paragraph{Conditional offloading in cryptodev}
%Note this is still under heavy criticism. If the hardware lags so far behind, why the relatively good score in openvpn?
The figure \ref{fig:openssl-benchmark-cbc} clearly shows a threshold on the packet size from which the hardware has a clear advantage, and below which the user mode software implementation is to go for.
%TODO include a listing to where it could be a good place to branch on the packet size. I already found it.
%TODO at least add a pseudo-code to illustrate this shit.
Using this breakeven point, one could patch the cryptodev cryptodev engine of OpenSSL to branch on the packet size, using the hardware if the packet is large enough, or fallback on the default software implementation otherwise. %openssl-1.0.2a/crypto/engine/eng_cryptodev.c
It would probably not be as trivial and beneficial as it may sound:
\begin{itemize}
	\item the encryption contexts would need to be synchronized between the hardware and the software;
	% \item we would already suffer from some context switches when sending the information to the cryptodev driver; %No, we are still in the openssl engine.
	% \item if we always branch to the fallback functions, the time overhead due to crytodev would lower the performance; %No, we don't leave the user space, the overhead is minimum.
	\item as the breakeven point is around 1024kB, the performance for a network application would be very close to those of a full software implementation, knowing that the ethernet frame size, the MTU, is set by default at 1500 bytes.
\end{itemize}

>>>>

\paragraph{Conditional offloading in cryptodev}
%Note this is still under heavy criticism. If the hardware lags so far behind, why the relatively good score in openvpn?
The figure \ref{fig:openssl-speed} clearly shows a threshold on the packet size from which the hardware has a clear advantage, and below which the user mode software implementation is to go for.
%TODO include a listing to where it could be a good place to branch on the packet size. I already found it.
%TODO at least add a pseudo-code to illustrate this shit.
Using this breakeven point, one could patch the cryptodev cryptodev engine of OpenSSL to branch on the packet size, using the hardware if the packet is large enough, or fallback on the default software implementation otherwise. %openssl-1.0.2a/crypto/engine/eng_cryptodev.c
It would probably not be as trivial and beneficial as it may sound:
\begin{itemize}
	\item the encryption contexts would need to be synchronized between the hardware and the software;
	% \item we would already suffer from some context switches when sending the information to the cryptodev driver; %No, we are still in the openssl engine.
	% \item if we always branch to the fallback functions, the time overhead due to crytodev would lower the performance; %No, we don't leave the user space, the overhead is minimum.
	\item as the breakeven point is around 1024kB, the performance for a network application would be very close to those of a full software implementation, knowing that the ethernet frame size, the MTU, is set by default at 1500 bytes.
\end{itemize}

### 2015-05-26 graphs/ipsec-ftp-bench.tex
Results

\addplot[style={black,fill=brown,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 14.68)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
        (aes256gcm, 0)
    };
    \label{inside tunnel}%"it" for "in tunnel"
\legend{software, ba411e, out-of-tunnel, inside tunnel}
\end{axis}

@@

\addplot[style={black,fill=RawSienna,mark=none}]
    coordinates {
        (none:none, 10.21)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
        (aes256gcm, 0)
    };
    \label{it-tp}
\legend{}
\end{axis}

>>>>

\addplot[style={black,fill=NavyBlue,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 14.68)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
        (aes256gcm, 0)
    };
    \label{inside tunnel}%"it" for "in tunnel"
\legend{software, ba411e, out-of-tunnel, inside tunnel}
\end{axis}

@@

\addplot[style={black,fill=MidnightBlue,mark=none}]
    coordinates {
        (none:none, 10.21)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
        (aes256gcm, 0)
    };
    \label{it-tp}
\legend{}
\end{axis}

### 2015-05-26 graphs/openssh-bench.tex
Results

%%%%%%%%%%%%%%%%%%%%%%%%
% throughput
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        title = {file transfer over SSH tunnel},
        width  = 0.95*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar=10pt,
        bar width=8pt,
        enlarge x limits={abs=1},
        ymajorgrids = true,
        ylabel = {Throughtput [MB/s]},
        xlabel = {},
        ymin=0, ymax=12,
        symbolic x coords={oot, none:none, aes256cbc:none, aes256cbc:sha256},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.25)}, anchor=north,legend columns=2},
        enlarge x limits=0.1,
    ]

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%
% throughput
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        title = {File transfer over an SSH tunnel},
        width  = 0.95*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar=10pt,
        bar width=8pt,
        enlarge x limits={abs=1},
        ymajorgrids = true,
        ylabel = {Throughtput [MB/s]},
        xlabel = {},
        ymin=0, ymax=12,
        symbolic x coords={oot, none:none, aes256cbc:none, aes256cbc:sha256},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.25)}, anchor=north,legend columns=2},
        enlarge x limits=0.1,
    ]

### 2015-05-26 graphs/openvpn-ftp-bench.tex
Results

\addplot[style={black,fill=brown,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 42.60)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
    };
    \label{inside tunnel}%"it" for "in tunnel"
\legend{software, ba411e, out-of-tunnel, inside tunnel}
\end{axis}

%%%%%%%%%%%%%%%%%%%%%%%%
% throughput
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        title = {FTP transfer inside OpenVPN tunnel},
        width  = 0.95*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar=10pt,
        bar width=8pt,
        enlarge x limits={abs=1},
        ymajorgrids = true,
        ylabel = {Throughtput [MB/s]},
        xlabel = {},
        ymin=0, ymax=12,
        symbolic x coords={oot, none:none, aes256cbc:none, aes256cbc:sha256},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.25)}, anchor=north,legend columns=2},
        enlarge x limits=0.1,
    ]

@@

\addplot[style={black,fill=RawSienna,mark=none}]
    coordinates {
        (none:none, 5.18)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
    };
    \label{it-tp}
\legend{}
\end{axis}

>>>>

\addplot[style={black,fill=NavyBlue,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 42.60)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
    };
    \label{inside tunnel}%"it" for "in tunnel"
\legend{software, ba411e, out-of-tunnel, inside tunnel}
\end{axis}

%%%%%%%%%%%%%%%%%%%%%%%%
% throughput
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        title = {FTP transfer inside an OpenVPN tunnel},
        width  = 0.95*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar=10pt,
        bar width=8pt,
        enlarge x limits={abs=1},
        ymajorgrids = true,
        ylabel = {Throughtput [MB/s]},
        xlabel = {},
        ymin=0, ymax=12,
        symbolic x coords={oot, none:none, aes256cbc:none, aes256cbc:sha256},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.25)}, anchor=north,legend columns=2},
        enlarge x limits=0.1,
    ]

@@

\addplot[style={black,fill=MidnightBlue,mark=none}]
    coordinates {
        (none:none, 5.18)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
    };
    \label{it-tp}
\legend{}
\end{axis}

### 2015-05-26 results.tex
Results

@@

\noindent Firstly, is the impact of OpenVPN's operation on the delay: when no encryption nor authentication is used, the exchange is 33\% to 53\% longer. This is simply because the packet has to go through a virtual interface, hence even without the security time overhead, the latency is much longer.

@@

\noindent In this case, the time overhead imposed by IPsec is not larger than 9\%, corroborating the results of~\citet{Xenakis20063225}.

@@

\begin{figure}[ht]
\input{graphs/ipsec-ping-benchmark}
\caption{IPsec: ping average response time}{for different packet sizes using IPsec. For each packet size, 1000 requests were flooded to the board, that is \textit{"outputs packets as fast as they come back or one hundred times per second, whichever is more"}, according to the \texttt{ping} command manual.} %TODO move the end of the caption to the "test protocol" chapter.
\label{fig:ping-benchmark-ipsec}
\end{figure}

@@

\noindent The first main difference is the overhead imposed by each method: 53\% for OpenVPN and only 9\% for IPsec.
Even if those configurations are not realistic, it puts forward the advantage of directly working in the kernel as does IPsec.

@@

\subsection{openSSH}

\begin{figure}[ht]
\input{graphs/openssh-bench}
\caption{file transfer over an SSH tunnel. The background stripped bars are the CPU usage.}{}
\label{fig:openssh-bench}
\end{figure}

\subsection{OpenVPN}

\begin{figure}[ht]
\input{graphs/openvpn-ftp-bench}
\caption{FTP file transfer over an OpenVPN tunnel. The background stripped bars are the CPU usage.}{}
\label{fig:openvpn-ftp-bench}
\end{figure}

We can see that adding a MAC computation aside the encryption merely lowers the performance when using the hardware.
Even though OpenSSL uses here an hihgly ARM-optimzed assembly implementation of SHA-256, it shows that the bottleneck is on the hardware side.

Indeed,

\subsection{IPsec}

\begin{figure}[ht]
\input{graphs/ipsec-ftp-bench}
\caption{FTP file transfer over an IPsec tunnel. The background stripped bars are the CPU usage.}{}
\label{fig:ipsec-ftp-bench}
\end{figure}

Some results have to be put into perspective with the fact that the implementation of SHA-256 is entirely C-based.
A more recent one using assembly instructions optimized for the NEON SIMD instruction set of the ARMv7 core could be used and would most probably yield better results.
The CPU usage of the software implementation would drop -- even if not significantly -- as for the hardware, it would be less limited by the software MAC counter part, and if the CPU usage could stay at the same level, we could expect a better throughput.

The GCM performance presented clearly shows a drop of throughput and an increase of CPU usage, illustrating the fact that those operations are hard on the software.
With an hardware offload, we could expect not only a drastic drop of the CPU usage, but an increase of throughput as well, since it's CPU-limited in those results.
Note that they are achieved usinf a C-based implementation of galois-field multiplications.
As we saw in chapter~\ref{chap:theory}, modern processor designers tend to add specialized instruction set aimed at AES-GCM enhancement.
Should further tests be conducted concerning IPsec paired with GCM, it would be wise to compare with an assembly implementation exploiting ARM NEON instruction set.
Some are being developed~\cite{Conrado2013,Danilo2013}, but none have been committed to the Linux kernel repository yet.

\subsection{Comparison}

\begin{figure}[ht]
\input{graphs/ftp-bench-comparison}
\caption{Comparison of file transfer methods}{The background stripped bars are the CPU usage.}
\label{fig:ftp-bench-comparison}
\end{figure}

>>>>

This chapter gather the results of the use cases presented in chapter~\ref{chap:test-protocol}.
Three use cases are considered: establishing TLS connections, exchanging ICMP requests to compute the latency and transfering a file over a secure channel.
Depending on the case, several schemes are analysed: DHE-RSA, AES-CBC, AES-GCM and SHA-256.

It is to be noted that all the CPU usage values are taken from the single one loaded core.
Although the platform is dual-core, all applications are used single threaded, either by design, such as OpenVPN, or by choice to obtain comparable results.

@@

Firstly, is the impact of OpenVPN's operation on the delay: when no encryption nor authentication is used, the exchange is 33\% to 53\% longer. This is simply because the packet has to go through a virtual interface, hence even without the security time overhead, the latency is much longer.

@@

\begin{figure}[ht]
\input{graphs/ipsec-ping-benchmark}
\caption{IPsec: ping average response time}{for different packet sizes.}
\label{fig:ping-benchmark-ipsec}
\end{figure}

In this case, the time overhead imposed by IPsec is not larger than 9\%, corroborating the results of~\citet{Xenakis20063225}.

@@

@@

The first main difference is the overhead imposed by each method: 53\% for OpenVPN and only 9\% for IPsec.
Even if those configurations are not realistic, it puts forward the advantage of directly working in the kernel as does IPsec.

@@

\subsection{OpenSSH}
The figure~\ref{fig:openssh-bench} shows the results for a file transfer over an SSH tunnel.
For this application, there is no none:none couple case inside the tunnel, as it would be the same as a transfer outside the tunnel.

\begin{figure}[ht]
\input{graphs/openssh-bench}
\caption{File transfer over an SSH tunnel.}{The background stripped bars are the CPU usage.}
\label{fig:openssh-bench}
\end{figure}

When only encryption is used, both implementation performs almost as well as outside the tunnel, but the software already saturates the CPU, whilst the hardware only uses half the same ressources.
About 41\% of the operations using the CPU when accelarating with the hardware involve interuptions and kernel memory management.

Adding the authentication makes the software performance drop by 25\%, as the CPU was already saturated without those extra operations.
The hardware is also consistent, but as there were some ressources available, the throughput is merely affected, even if the MAC is not offloaded in hardware.

\subsection{OpenVPN}
The figure~\ref{fig:openvpn-ftp-bench} shows the performance of the file transfer over an OpenVPN tunnel.

\begin{figure}[ht]
\input{graphs/openvpn-ftp-bench}
\caption{FTP file transfer over an OpenVPN tunnel.}{The results inside the tunnel with no encryption nor authentication highlights the overhead OpenVPN adds to the transfer. From there, the throughput only change by a small extent. The poor hardware results also shows the heaviness of OpenVPN on hardware offloading. The background stripped bars are the CPU usage.}
\label{fig:openvpn-ftp-bench}
\end{figure}

As it was already the case for the ping benchmark, the overhead imposed by the manipulation of OpenVPN is extremely heavy: the CPU usage jumps from 7.16\% to 42.60\%, and the throughput drops by 55\%, from 11.39MB/s to 5.18MB/s.

\noindent Surprisingly enough, encrypting the data only changes the throughput by 8\%, but the CPU usage is almost doubled.
A fair interpretation would be that the encryption is no the bottleneck, the transfer bewteen the user and kernel mode, as well as the fragmentation, are.
However for the hardware, the performance collapse to 3.35MB/s.

Adding a MAC computation aside the encryption lowers the performance in software and hardware by respectively 20\% and 15\%, for the same CPU usage as without authentication.

\subsection{IPsec}
The figure~\ref{fig:ipsec-ftp-bench} gathers the results of the file transfer over the IPsec tunnel.
\begin{figure}[ht]
\input{graphs/ipsec-ftp-bench}
\caption{FTP file transfer over an IPsec tunnel.}{The advantage of the hardware is a lower CPU usage, at the cost of a few percents lower throughput. The GCM mode is also tested to discuss its implementation in hardware. The background stripped bars are the CPU usage.}
\label{fig:ipsec-ftp-bench}
\end{figure}

In this implementation, the IPsec overhead is only of 10\%, close to the 9\% of the ping overhead, and the CPU usage only double to reach 14.68\%.

When encrypting the data, the CPU utilization explodes in the case of the software implementation, but stays at the same level for the hardware.
As for the throughput, it lowers respectively by 14\% and 17\%.
What could have been expected for the hardware implementation is a more stable throughput, but an increase of CPU usage for the few operations undergone by the BA411E driver, especially considering the fact that it uses active polling on the hardware.
Thus, even if the hardware is the bottleneck, the CPU usage should be higher.

\noindent An explanation could be that the operating system prevent the driver to monopolizing the CPU with its polling and preempt it regularly, effectively lowering the ressource usage, but also limiting the performance.

Adding the authentication yields expected results for : an increase of CPU usage and lower throughput.
Both implementation lose 2MB/s and a CPU usage increase of 20\% and 16\%.

In both cases, the hardware implementation exhibited slightly lower performance, but a CPU usage three to four times lower.
On embedded platforms, this is as important as the raw performance, because lower CPU usage means underclocked CPU, resulting in a lower power consumption.

The softaware GCM results allows to open a discussion on this mode used in conjunction with IPsec.
The GCM performance presented clearly shows a drop of throughput and an increase of CPU usage, illustrating the fact that those operations are hard on the software.
With an hardware offload, we could expect not only a drastic drop of the CPU usage, but an increase of throughput as well, since it's CPU-limited in those results.
Note that they are achieved using a C-based implementation of galois-field multiplications.
As we saw in chapter~\ref{chap:theory}, modern processor designers tend to add specialized instruction sets aimed at AES-GCM enhancement.
Should further tests be conducted concerning IPsec paired with GCM, it would be wise to compare with an assembly implementation exploiting ARM NEON instruction set.
Some are being developed~\cite{Conrado2013,Danilo2013}, but none have been committed to the Linux kernel repository yet.

\subsection{Comparison}
The figure~\ref{fig:ftp-bench-comparison} summarizes the file transfer performance for all the implementations, using the most realistic configuration tested: AES-256-CBC with SHA-256.
\begin{figure}[ht]
\input{graphs/ftp-bench-comparison}
\caption{Comparison of file transfer methods}{Out of the three methods, OpenSSH is to be prefered is the raw performance is the objective, but IPsec offers an higher effeciency. At the same time, OpenVPN is worse on all fronts. The background stripped bars are the CPU usage.}
\label{fig:ftp-bench-comparison}
\end{figure}

In this work, we have two objectives: the raw performance and the CPU usage.
The first objetive is fulfilled by OpenSSH; even if it is a user-space application that needs to transfer the data through the kernel in order to use the hardware, it still outperforms the software by 27\%.

\noindent The second objective is best fulfilled by IPsec with an efficiency of 0.34MB/s/\% of CPU for the hardware, and 0.09MB/s/\% of CPU for the software.

OpenVPN is however losing on all fronts. Not only it has the lowest throughput, but the CPU usage of the hardware is out the roof, especially when compared to the performance.

\noindent The table~\ref{tab:ftp-fragmentation} shows to what size are fragmented the packets by OpenVPN and IPsec.
A look at the figure~\ref{fig:openssl-speed} shows that fragmenting at small size is a bad idea, especially when the hardware is reached from the user space.
OpenVPN sends one third of its packets at sizes that are highly ineffective, whilst IPsec manages it better and limit the small packets to less than 10\%.

\begin{table}
\center
\begin{tabular}{l|c|c|} \cline{2-3}
 & Length & Frequency \\ \hline
\multicolumn{1}{|l|}{\multirow{3}{*}{OpenVPN}} & 80 -- 159 & 33.04\% \\ \cline{2-3}
\multicolumn{1}{|l|}{} & 1280 -- 2559 & 66.90\% \\ \cline{2-3}
\multicolumn{1}{|l|}{} & other & 0.06\% \\ \hline
\multicolumn{1}{|l|}{\multirow{3}{*}{IPsec}} & 80 -- 159 & 9.42\% \\ \cline{2-3}
\multicolumn{1}{|l|}{} & 1280 -- 2559 & 89.84\% \\ \cline{2-3}
\multicolumn{1}{|l|}{} & other & 0.74\% \\ \hline
\end{tabular}
\caption{Packet size frequency for an FTP transfer}{OpenVPN is less effective in its fragmentation, producing more packets of a smaller size, which perform worse on hardware.}
\label{tab:ftp-fragmentation}
\end{table}

\begin{figure}[ht]
\input{graphs/openssl-speed}
\caption{OpenSSL benchmark for AES-256-CBC}{compare a run of \texttt{openssl speed} for three implementation of AES: in software by OpenSSL (soft), in software by the standard Linux kernel module (soft-drv) and in hardware (hard). The last two have to go through cryptodev in order to be reached from the user space.}
\label{fig:openssl-speed}
\end{figure}

Some results have to be put into perspective with the fact that the implementation of SHA-256 is entirely C-based.
A more recent one using assembly instructions optimized for the NEON SIMD instruction set of the ARMv7 core could be used and would most probably yield better results.
The CPU usage of the software implementation would lower -- even if not significantly -- as for the hardware, it would be less limited by the software MAC counter part, and if the CPU usage would stay at the same level, we could expect a better throughput.
%TODO table comparing the packet size (wireshark)

### 2015-05-26 test-proto.tex
moar theoretical works
Results

\chapter{Test protocol}

@@

@@

@@

>>>>

\chapter{Test protocol}\label{chap:test-protocol}

@@

As this experiment can be very unstable and vastly depends on the operating system scheduling, each test case has been repeated five times to ensure stable results.

@@

The security parameter tested is a standard TLS-DHE-RSA, hence forcing the peers the renogociate a new shared secret and ephemeral Diffie-Hellman parameters at each connection attempt.

\section{Response time -- latency}
This use case exchanges ICMP request of various sizes via the \texttt{ping} command.
The initiating peer sends an ICMP echo request to the remote peer, which then answers with an ICMP echo reply.

\noindent For each packet size, 1000 requests were flooded to the board, that is \textit{"outputs packets as fast as they come back or one hundred times per second, whichever is more"}, according to the \texttt{ping} command manual.

The following loop shows the options used for the test as well as the payload sizes:
\begin{lstlisting}[language=bash]
for i in 56 1000 8000 16000; do
	sudo ping -f -s ${i} -c 1000 150.158.232.241
done
\end{lstlisting}

@@

We will also use two types of drivers for the BA411E: an interuption-based for OpenVPN and OpenSSH, and a polling-based for IPsec.

All the tests will be conducted with the ESP protocol in tunnel mode, so that we have the worst case scenario; as AH imposes a smaller overhead, the perfomance could only be better.

For OpenSSH, the commmand \texttt{scp} will be used to transfer the data securely over an SSH tunnel.
As for OpenVPN and IPsec, a tunnel will be established beforehand, and then the simple \texttt{ftp} command will allow the client to fetch the file on the server.
